Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2020-02-18 xrb/submit/paper.tex
Added paragraph with flame speed comparison

The computational expense of the problem necessitated several approximations
to make this set of simulations \citep{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal, Cavecchi2019}
tractable. The first was the assumption of vertical hydrostatic equilibrium, which
allowed them to filter out vertically propagating sound waves that would otherwise
have limited the timestep. This enabled them to reach the simulation times (on the
order of seconds) needed to capture the rise of the burst, but they could not model
vortical fluid motions driven by effects such as turbulence and convection. These can
facilitate mixing and help to accelerate the propagation of the flame front. They also
included only the triple-$\alpha$ process in their implementation of nuclear burning,
and use a fixed, constant opacity across the entire domain for each simulation run (although
they explore the dependence of the flame propagation speed on this opacity value).
Our simulations relax the assumption of vertical hydrostatic equilibrium, and implement
a more complete set of nuclear reactions and opacities. However, this comes at the expense
of a different set of approximations.

@@

>>>>

The computational expense of the problem necessitated several approximations
to make this set of simulations \citep{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal, Cavecchi2019}
tractable. The first was the assumption of vertical hydrostatic equilibrium, which
allowed them to filter out vertically propagating sound waves that would otherwise
have limited the timestep. This enabled them to reach the simulation times (on the
order of seconds) needed to capture the rise of the burst, but they could not model
vortical fluid motions driven by effects such as turbulence and convection. These can
facilitate mixing and help to accelerate the propagation of the flame front. They also
included only the triple-$\alpha$ process in their implementation of nuclear burning,
and use a fixed, constant opacity across the entire domain for each simulation run (although
they explore the dependence of the flame propagation speed on this opacity value).
Our simulations relax the assumption of vertical hydrostatic equilibrium, and implement
a more complete set of nuclear reactions and opacities. However, as discussed later, this
comes at the expense of a different set of approximations.

@@

\citet{cavecchi:2013} ran simulations at rotation rates of $450$ Hz exploring a range of opacities.
Accounting for the boosting, we would expect to see our 10/10 flame propagate at $\sim 2.25$ times
the speed of these $450$ Hz runs, but instead the speeds range from $4.5$ to $7$ times those reported in
\citet{cavecchi:2013}. The additional boost may come from cyclic fluid motions that would have been
disallowed in the $450$ Hz runs by the enforcement of vertical hydrostatic equilibrium (we already saw
that hydrodynamic effects can accelerate the burning front in our comparison with the 1D flames). The
inclusion of reaction chains beyond may also increase energy generation and consequently the
propagation speed.
